---
title: "Exercise 2 - Data Preparation"
author: Ryan Wesslen
date: Sept 27, 2016
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

### Objective: Analyze the major presidential candidates Tweets

Our goal is to analyze the rhetoric (if you can call it that) of the Tweets for four major political candidates for president: Hillary Clinton, Bernie Sanders, Donald Trump and Ted Cruz.

The dataset we'll use is includes all of the candidates' Tweets between June 2015 to June 2016 and includes about 19,000 Tweets (that's about 8k Tweets per candidate!).

As we mentioned in yesterday's session, there are two ways to run this RMarkdown file: as automatically as a Knit (e.g. to produce a HTML file) or in manually chunks.

If you run it in chunks (as we will in this tutorial), you will need to set your working directory and remove one of the "." from the `read.csv()` and `source()` functions. If you are running it as Knit file, you can leave as is.

#### Step 1: Read in the data. 

Like yesterday, our first steps include setting our working directory (if running as chunks) and loading in our dataset.

```{r}
#set your personal working directory if you're running as chunks
#setwd("~/Dropbox/fall-2016-pm-twitter-text/")

#remove one of the "." if you are running as chunks
raw.tweets <- read.csv("../datasets/pres-tweets.csv", stringsAsFactors = F)
source('../functions.R')
```

#### Step 2: Explore the dataset.

Let's explore our dataset. You can either open the dataset or run the `str()` function.

```{r}
str(raw.tweets)
```

We have 19,103 Tweets, averaging about 4-5k Tweets per candidate.

#### Step 3: Explore the data

We'll use the `dplyr` package to explore aggregate counts in the dataset.

```{r}
library(dplyr)

agg_tweets <- raw.tweets %>% 
  group_by(displayName) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(Count))

agg_tweets
```

Further, let's first plot the Tweet count per candidate by day.

```{r}
library(ggplot2)

raw.tweets$date <- as.Date(raw.tweets$estTime, "%Y-%m-%d")
date_tweets <- raw.tweets %>% 
  group_by(date, displayName) %>% 
  summarise(Count = n())

ggplot(date_tweets, aes(date, Count, colour = displayName, group = displayName)) + 
  geom_line()
```

Unfortunately this is a bit noisy. Let's instead go with month. We'll use the `zoo` package to extract the year-month for all Tweets.

```{r}
temp <- zoo::as.yearmon(raw.tweets$date)

raw.tweets$yearmonth <- paste(format(temp, "%Y"),format(temp, "%m"))

date_tweets <- raw.tweets %>% 
  group_by(yearmonth, displayName) %>% 
  summarise(Count = n())

ggplot(date_tweets, aes(yearmonth, Count, colour = displayName, group = displayName)) + 
  geom_line()
```

This is much clearer that Donald Trump is the heaviest Tweeter, with the largest number of Tweets in July 2015 and Oct 2015 with nearly **800 Tweets** per month or about 27 Tweets per day!

#### Step 4: Identify the most common Hashtags and Handles

Find the most common hashtags with the function `getCommonHashtags()`. Like the `timePlot()` function, this function is in the functions.R file. 

For the function, use the Tweet text (body) as the input. 

```{r}
hashtags <- getCommonHashtags(raw.tweets$body)

head(hashtags, 25)
```

What are the top 25 hashtags? Which are Panther related?

Repeat the same exercise but with the function `getCommonHandles()` to find the most common handles.

```{r}
handles <- getCommonHandles(raw.tweets$body)

head(handles, 25)
```

#### Step 5: Favorite and Retweet Count

This data also includes the number of times each Tweet was favorited (liked) and retweeted. This is a measurement of the response on social media by all users.

Let's first explore the distribution of Favorited by each candidate using a box and whisker plot.

```{r}
library(ggplot2)
ggplot(raw.tweets, aes(displayName, favoritesCount)) + 
  geom_boxplot() +
  ylab("Favorited Tweet Count (Likes)") 
```

Unfortunately, this is distorted because there's a huge tail. Let's use change the Y-axis to log-based.

```{r}
ggplot(raw.tweets, aes(displayName, favoritesCount)) + 
  geom_boxplot() +
  ylab("Favorited (Likes) Tweet Count - Log Scale") + scale_y_log10()
```

Ah - so we see that Donald Trump on average has a larger response than the other candidates during this period. Interestingly, Bernie has (on average) the second most number Tweets per candidate.

Let's also plot the number of Retweets.

```{r}
ggplot(raw.tweets, aes(displayName, retweetCount)) + 
  geom_boxplot() +
  ylab("Retweet Tweet Count - Log Scale") + scale_y_log10()
```

If we want to get the actual summary statistics, we can use the `psych` package:

```{r}
library(psych)
describeBy(raw.tweets$favoritesCount, group = raw.tweets$displayName)
describeBy(raw.tweets$retweetCount, group = raw.tweets$displayName)
```

Let's now get the top 10 Tweets for each candidate by the largest number of "Favorites":

```{r}
candidates <- c("Bernie Sanders","Hillary Clinton","Donald J. Trump", "Ted Cruz")

for (i in 1:4){
  ctweets <- raw.tweets[raw.tweets$displayName == candidates[i],]
  row <- order(-ctweets$favoritesCount)
  print(candidates[i], row.names = FALSE)
  print(ctweets[row[1:10],c(4,2,5,7)], row.names = FALSE)
}
dev.off()
```

FYI - there are much better ways to display results than the `print` function. Consider using the `DT` or `gridExtra` for future tables.

Let's combine our analysis from the time series plots and the Favorite counts to create plots for the candidates.

```{r}
#Bernie
ctweets <- raw.tweets[raw.tweets$displayName == "Bernie Sanders",]
ggplot(raw.tweets, aes(yearmonth, favoritesCount)) + 
  geom_boxplot() + 
  ggtitle("Bernie Sanders") + 
  ylab("Favorited (Likes) Tweet Count - Log Scale") + 
  scale_y_log10()

#Hillary
ctweets <- raw.tweets[raw.tweets$displayName == "Hillary Clinton",]
ggplot(raw.tweets, aes(yearmonth, favoritesCount)) + 
  geom_boxplot() + 
  ggtitle("Hillary Clinton") + 
  ylab("Favorited (Likes) Tweet Count - Log Scale") + 
  scale_y_log10()

#Trump
ctweets <- raw.tweets[raw.tweets$displayName == "Donald J. Trump",]
ggplot(raw.tweets, aes(yearmonth, favoritesCount)) + 
  geom_boxplot() + 
  ggtitle("Donald J. Trump") + 
  ylab("Favorited (Likes) Tweet Count - Log Scale") + 
  scale_y_log10()

#Cruz
ctweets <- raw.tweets[raw.tweets$displayName == "Ted Cruz",]
ggplot(raw.tweets, aes(yearmonth, favoritesCount)) + 
  geom_boxplot() + 
  ggtitle("Ted Cruz") + 
  ylab("Favorited (Likes) Tweet Count - Log Scale") + 
  scale_y_log10()
```